---
title: "mlflx data preparation"
author: "Beni"
date: "1/5/2021"
output: html_document
---


```{r setup, include=FALSE}
library(tidyverse)
library(ingestr)
library(rbeni)
```

## Site selection

Site selection was done based on:

1. Site where fLUE method worked. I.e., site belongs to clusters cDD, cGR, cLS, or cNA, as described in Stocker et al. (2018) *New Phytologist*.

*ADDITIONAL SELECTION STEP BELOW NOT APPLIED*
2. Site was among the subset of homogenous sites, selected by Manuela Balzarolo. All sites with a homogenous surrounding are listed in file `~/data/FLUXNET-2015_Tier1/meta/fluxnet_quality_check_homogenous_OK_PRI_mbalzarolo.csv`. 

Reproduce it as follows:

Get sites where fLUE data is available
```{r}
df_sites <- read_csv("~/data/flue/flue_stocker18nphyt.csv")
df_sites <- dplyr::select(df_sites, site, cluster) %>% unique()
```

Add meta information for sites
```{r}
df_sites <- df_sites %>% 
  rename(sitename = site) %>% 
  left_join( siteinfo_fluxnet2015, by="sitename" )
```

<!-- Get list of sites with homogenous surrounding and subset fLUE sites based on that list -->
<!-- ```{r} -->
<!-- df_homo <- read_csv("~/data/FLUXNET-2015_Tier1/meta/fluxnet_quality_check_homogenous_OK_PRI_mbalzarolo.csv") -->

<!-- df_flue_sites <- df_flue_sites %>%  -->
<!--   mutate(homogenous_mbalzarolo = sitename %in% df_homo$sitename) -->

<!-- df_sub_homo <- df_flue_sites %>%  -->
<!--   dplyr::filter( homogenous_mbalzarolo & cluster %in% c("cGR", "cDD", "cLS", "cNA") ) -->
<!-- ``` -->

Write to file
```{r}
if (!dir.exists("./data")) system("mkdir data")
write_csv(df_sites, path = "./data/df_sites.csv")
```

Show sites on a map.
```{r}
plot_map_simpl() +
  geom_point(data = df_sites, aes(x = lon, y = lat), color = "red")
```

## FLUXNET data

### Read data

```{r message=FALSE}
read_dd_bysite <- function(site){
  dir <- "~/data/FLUXNET-2015_Tier1/20191024/DD"
  filn <- list.files(dir, pattern = paste0("FLX_", site, "_FLUXNET2015_FULLSET_DD_"), full.names = TRUE)
  read_csv(filn)
}
list_df <- purrr::map(as.list(df_sites$sitename), ~read_dd_bysite(.))
names(list_df) <- df_sites$sitename  # this makes it a named list
```

### Clean data

DD-specific
```{r}
## function definition
clean_fluxnet_dd <- function(df){
  
  df %>%

    ## select only the variables we're interested in
    dplyr::select(starts_with("TIMESTAMP"),
           ends_with("_F"),
           USTAR,
           CO2_F_MDS,
           # ends_with("QC"),
           -contains("JSB"),
           starts_with("SWC"),
           GPP_NT_VUT_REF,
           NEE_VUT_REF_QC
           ) %>%

    ## convert to a nice date object
    mutate(TIMESTAMP = lubridate::ymd(TIMESTAMP)) %>%

    ## set bad data to NA for multiple variables
    mutate(
      GPP_NT_VUT_REF = ifelse(NEE_VUT_REF_QC > 0.8, GPP_NT_VUT_REF, NA)
      ) %>%
    
    ## set all -9999 to NA
    na_if(-9999) %>%

    ## drop QC variables (no longer needed), except NEE_VUT_REF_QC
    dplyr::select(-ends_with("_QC"), NEE_VUT_REF_QC)
}

list_df <- purrr::map(list_df, ~clean_fluxnet_dd(.))
list_df[1]
```

Make it a single flat dataframe. Has now shape 265,175 x 19.
```{r}
ddf <- list_df %>% 
  bind_rows()
```



<!-- HH-specific -->
<!-- ```{r} -->
<!-- clean_fluxnet_hh <- function(df){ -->

<!--   df <- df %>%  -->

<!--     ## select only the variables we're interested in -->
<!--     dplyr::select( -->
<!--       starts_with("TIMESTAMP"), -->
<!--       ends_with("_F"), -->
<!--       CO2_F_MDS, -->
<!--       PPFD_IN,  -->
<!--       GPP_NT_VUT_REF, -->
<!--       starts_with("SWC_F_MDS"), -->
<!--       NEE_VUT_REF_QC, -->
<!--       USTAR, VPD, RH, -->
<!--       ends_with("QC"), -->
<!--       -contains("JSB"), -->
<!--       NIGHT -->
<!--       ) %>%  -->

<!--     ## convert to nice time object -->
<!--     mutate_at(vars(starts_with("TIMESTAMP_")), ymd_hm) %>%  -->

<!--     ## set bad data to NA for multiple variables -->
<!--     mutate( -->
<!--       GPP_NT_VUT_REF = ifelse(NEE_VUT_REF_QC %in% c(0,1), GPP_NT_VUT_REF, NA), -->
<!--       TA_F = ifelse(TA_F_QC %in% c(0,1), TA_F, NA), -->
<!--       SW_IN_F = ifelse(SW_IN_F_QC %in% c(0,1), SW_IN_F, NA), -->
<!--       LW_IN_F = ifelse(LW_IN_F_QC %in% c(0,1,2), LW_IN_F, NA),   # relaxing filter criterion -->
<!--       VPD_F = ifelse(VPD_F_QC %in% c(0,1), VPD_F, NA), -->
<!--       PA_F = ifelse(PA_F_QC %in% c(0,1,2), PA_F, NA),   # relaxing filter criterion -->
<!--       P_F = ifelse(P_F_QC %in% c(0,1,2), P_F, NA),   # relaxing filter criterion -->
<!--       WS_F = ifelse(WS_F_QC %in% c(0,1), WS_F, NA), -->
<!--       CO2_F_MDS = ifelse(CO2_F_MDS_QC %in% c(0,1), CO2_F_MDS, NA), -->
<!--       SWC_F_MDS_1 = ifelse(SWC_F_MDS_1_QC %in% c(0,1), SWC_F_MDS_1, NA), -->
<!--       SWC_F_MDS_2 = ifelse(SWC_F_MDS_2_QC %in% c(0,1), SWC_F_MDS_2, NA), -->
<!--       SWC_F_MDS_3 = ifelse(SWC_F_MDS_3_QC %in% c(0,1), SWC_F_MDS_3, NA), -->
<!--       SWC_F_MDS_4 = ifelse(SWC_F_MDS_4_QC %in% c(0,1), SWC_F_MDS_4, NA) -->
<!--       ) %>% -->

<!--     ## set all -9999 to NA -->
<!--     na_if(-9999) %>% -->

<!--     ## drop QC variables (no longer needed), except NEE_VUT_REF_QC -->
<!--     dplyr::select(-ends_with("_QC"), NEE_VUT_REF_QC) -->

<!--   return(df) -->
<!-- } -->
<!-- ``` -->




## MODIS data

This uses the ingestr package to download and interpolate MODIS time series.

```{r}
settings_modis <- get_settings_modis(
  bundle            = "modis_fpar",
  data_path         = "~/data/modis_subsets/",
  method_interpol   = "loess",
  keep              = TRUE,
  overwrite_raw     = FALSE,
  overwrite_interpol= TRUE
  )

df_modis_fpar <- ingest(
  df_sites %>% slice(1:2), 
  source = "modis",
  settings = settings_modis, 
  parallel = FALSE
  # ncores = 2
  )
```

Test plot.
```{r}
plot_fapar_ingestr_bysite(
  df_modis_fpar$data[[2]], 
  settings_modis)
```
